Parsing Postfix Log Files

### 2009-08-24 doc/logparser-introduction.tex
Provide a brief introduction to Postfix in the abstract and introduction, as instructed by Howard Hamilton.

The approach chosen to measure performance is to analyse the log files
produced by the \acronym{MTA} in use, Postfix, rather than modifying its
source code to generate statistics: this approach improves the chances of
other Postfix users testing, using, and benefiting from the software
developed for this project.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by Postfix: the
parser must be designed and implemented so that adding support for parsing
new inputs is a simple task, because the log lines to be parsed will change
over time.  This variety in log lines occurs for several reasons:

>>>>

The approach chosen to measure performance is to analyse the log files
produced by
Postfix,\footnote{\urlLastChecked{http://www.postfix.org/}{2009/04/21}.  An
introduction to Postfix is provided in \sectionref{postfix background}.} the
\acronym{MTA} used by the School of Computer Science and Statistics, rather
than modifying its source code to generate statistics: this approach
improves the chances of other Postfix users testing, using, and benefiting
from the software developed for this project.  The need arose for a parser
capable of dealing with the great number and variety of log lines produced
by Postfix: the parser must be designed and implemented so that adding
support for parsing new inputs is a simple task, because the log lines to
be parsed will change over time.  This variety in log lines occurs for
several reasons:

### 2009-08-24 doc/logparser.tex
Changes made to improve appearance after switching to one-and-a-half spacing.
Provide a brief introduction to Postfix in the abstract and introduction, as instructed by Howard Hamilton.

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix Log Files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

@@

    {\Huge Parsing Postfix Log Files}

@@

Parsing Postfix log files is much more difficult than it first appears, but
it \textit{is\/} possible to achieve a high degree of accuracy in
understanding Postfix log files, and subsequently in reconstructing the
actions taken by Postfix when processing mail delivery attempts.  This
thesis describes the creation of a parser for Postfix log files,
documenting the architecture developed for this project and the parser that
implements it, the difficulties encountered and the solutions developed.
The parser stores data gleaned from the log files in an SQL database;
future projects or programs could use the gathered data to optimise current
anti-spam measures, to produce statistics showing how effective those
measures are, or to provide a baseline to test new anti-spam measures
against.  The Postfix log file parser needs to be very precise and strict
when parsing, yet must allow users to easily adapt or extend it to parse
new log lines, without requiring that the user have an in-depth knowledge
and understanding of the parser's internal workings.  The newly developed
architecture is designed to make the process of parsing new inputs as
simple as possible, enabling users to trivially add new rules (to parse
variants of known inputs) and relatively easily add new actions (to process
a previously unknown category of inputs).  The parser implemented for this
project is evaluated on the criteria of efficiency and coverage of Postfix
log files, demonstrating that the conflicting goals of efficiency and
accuracy can be balanced, and that one need not be sacrificed to achieve
the other.

@@

\newpage

>>>>

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\newcommand{\mytitle}[0]{%
    Parsing Log Files Produced by the Postfix Mail Transfer Agent%
}
\hypersetup{
    pdftitle    = {\mytitle},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

@@

    {\Huge \mytitle{}}

@@

Postfix is a widely used Mail Transfer Agent, transferring hundreds of
millions of mails between senders and recipients every day.  Parsing log
files produced by Postfix is much more difficult than it first appears, but
it \textit{is\/} possible to achieve a high degree of accuracy in
understanding Postfix log files, and subsequently in reconstructing the
actions taken by Postfix when processing mail delivery attempts.  This
thesis describes the creation of a parser for Postfix log files,
documenting the architecture developed for this project and the parser that
implements it, the difficulties encountered and the solutions developed.
The parser stores data gleaned from the log files in an SQL database;
future projects or programs could use the gathered data to optimise current
anti-spam measures, to produce statistics showing how effective those
measures are, or to provide a baseline to test new anti-spam measures
against.  The Postfix log file parser needs to be very precise and strict
when parsing, yet must allow users to easily adapt or extend it to parse
new log lines, without requiring that the user have an in-depth knowledge
and understanding of the parser's internal workings.  The newly developed
architecture is designed to make the process of parsing new inputs as
simple as possible, enabling users to trivially add new rules (to parse
variants of known inputs) and relatively easily add new actions (to process
a previously unknown category of inputs).  The parser implemented for this
project is evaluated on the criteria of efficiency and coverage of Postfix
log files, demonstrating that the conflicting goals of efficiency and
accuracy can be balanced, and that one need not be sacrificed to achieve
the other.

@@
